Non-word repetition in Yélî Dnye

### 2020-09-08 manuscript.Rmd
test IPA font with linguisticsdown

lang              : "english"
class             : "man"
output            :
  papaja::apa6_pdf:
    latex_engine: xelatex
---

@@

The Yélî sound system, much like its baroque grammatical system in general [@levinsoninprepydgrammar], is unlike any other in the region, with 90 contrastive segments, including at least two contrasts not yet been documented elsewhere (labial-coronal double-articulations with dental vs. post-alveolar coronal placement in both oral and nasal stops) [@levinsoninprepydgrammar]. With only four primary places of articulation (bilabial, alveolar, post-alveolar, and velar) and no voicing contrasts, the phonological inventory is remarkably packed with acoustically similar segments. The core oral stop set includes both singleton (/p/, /t/, /ṭ/, and /k/) and doubly-articulated (/tp/, /ṭp/, /kp/) segments, with full nasal equivalents (/m/, /n/, /ṇ/, /ŋ/, /nm/, /ṇm/, /ŋm/), and with a substantial portion of these able to be pre-nasalized or nasally released (/mp/, /nt/, /ṇṭ/, /ŋk/, /nmtp/, /ṇmṭp/, /ŋmkp/, /ṭṇ/, /kŋ/, /ṭpṇm/, /kpŋm/). Finally, a number of these can also be labialized or palatalized on the release, or both (/pʲ/, /tʃ/, /tpʲ/, /ṭʲ/, /ṭpʲ/, /kʲ/, /kpʲ/, /pʲʷ/, /mbʲ/, /ndʒ/, /nmdbʲ/, /ṇḍʲ/, /ṇmḍbʲ/, /mbʷ/, /ŋɡʷ/, /mbʲʷ/, /ṭṇʲ/, /ṭpṇmʲ/, /kŋʷ/, /mʲ/, /nʲ/, /nmʲ/, /ṇʲ/, /ṇmʲ/, /mʷ/, /ŋʷ/, mʲʷ/). Among the remaining consonants are small group of oral continuants (/w/, /j/, /ɣ/, /l/, /βʲ/, /lʲ/, /lβʲ/). Vowels in Yélî Dnye may be oral or nasal, short or long. The 10 oral vowel qualities (/i/, /ɯ/, /u/, /e/, /o/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/) can be produced as short or long vowels, with 7 of these able to appear as short and long nasal vowels as well (/i/, /u/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/). In total, Levinson [-@levinsoninprepydgrammar] then counts 90 distinctive segments in Yélî Dnye (93 when including consonants that are extremely rarely used).

>>>>

lang              : "english"
class             : "man"
output            :
  papaja::apa6_pdf:
    latex_engine: xelatex
    includes:
      in_header: header.tex
---

@@

The Yélî sound system, much like its baroque grammatical system in general [@levinsoninprepydgrammar], is unlike any other in the region, with 90 contrastive segments, including at least two contrasts not yet been documented elsewhere (labial-coronal double-articulations with dental vs. post-alveolar coronal placement in both oral and nasal stops) [@levinsoninprepydgrammar]. With only four primary places of articulation (bilabial, alveolar, post-alveolar, and velar) and no voicing contrasts, the phonological inventory is remarkably packed with acoustically similar segments. The core oral stop set includes both singleton (/p/, /t/, /ṭ/, and /k/) and doubly-articulated (/tp/, /ṭp/, /kp/) segments, with full nasal equivalents (/m/, /n/, /ṇ/, /ŋ/, /nm/, /ṇm/, /ŋm/), and with a substantial portion of these able to be pre-nasalized or nasally released (/mp/, /nt/, /ṇṭ/, /ŋk/, /nmtp/, /ṇmṭp/, /ŋmkp/, /ṭṇ/, /kŋ/, /ṭpṇm/, /kpŋm/). Finally, a number of these can also be labialized or palatalized on the release, or both (/pʲ/, /tʃ/, /tpʲ/, /ṭʲ/, /ṭpʲ/, /kʲ/, /kpʲ/, /pʲʷ/, /mbʲ/, /ndʒ/, /nmdbʲ/, /ṇḍʲ/, /ṇmḍbʲ/, /mbʷ/, /ŋɡʷ/, /mbʲʷ/, /ṭṇʲ/, /ṭpṇmʲ/, /kŋʷ/, /mʲ/, /nʲ/, /nmʲ/, /ṇʲ/, /ṇmʲ/, /mʷ/, /ŋʷ/, mʲʷ/). Among the remaining consonants are small group of oral continuants (/w/, /j/, /ɣ/, /l/, /βʲ/, /lʲ/, /lβʲ/). Vowels in Yélî Dnye may be oral or nasal, short or long. The 10 oral vowel qualities (/i/, /ɯ/, /u/, /e/, /o/, /ə/, /ɛ/, /`r linguisticsdown::cond_cmpl("ɔ")`/, /æ/, /ɑ/) can be produced as short or long vowels, with 7 of these able to appear as short and long nasal vowels as well (/i/, /u/, /ə/, /ɛ/, /`r linguisticsdown::cond_cmpl("ɔ")`/, /æ/, /ɑ/). In total, Levinson [-@levinsoninprepydgrammar] then counts 90 distinctive segments in Yélî Dnye (93 when including consonants that are extremely rarely used).

### 2020-09-08 manuscript.tex
test IPA font with linguisticsdown

\shorttitle{NWR in Yélî Dnye}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d'Etudes cognitives, ENS, EHESS, CNRS, PSL University\\\textsuperscript{2} Max Planck Institute for Psycholinguistics}
\keywords{phonology, non-word repetition, development\newline\indent Word count: ~xxx words}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\usepackage{lineno}

\linenumbers

@@

\begin{itemize}
\tightlist
\item
  look through this manuscript for bits of text to be written by
  \textbf{mc}
\item
  help with table 1 (formatting \& making it appear, because it's broken
  by the phonological representations)
\item
  do you think Figure 2 should be done with phonetic symbols? If yes,
  could you try to implement that?
\item
  double check demo -- probably start with the xlsx version that has the
  children's first names, date of birth, and date of test. I looked
  again at my local files (still haven't checked the labs' ones) and I'm
  surprised I have maternal education for more children than I have
  birth order for\ldots{} If you correct this, to generate the csv I
  also round the age, assign age.rounded to reported age (in years) when
  DOB is missing, remove first names, DOB, DOTest and age.precise before
  pushing to github
\item
  double check my frequency entries in segments.xlsx using the search
  function in \url{https://phoible.org/parameters}; take a look at
  \url{http://phoible.github.io/conventions/} in case you see something
  about double articulation (I didn't find tp or lBj)
\item
  double check phonological implementation of stimuli in stimuli.xlsx
\item
  for both of the above, if you make any changes, they need to be
  re-saved as UTF16 txt file, and then converted (see remember above)
\item
  double check also stim\_seg\_freq -- why are some items still with NA
  freq?
\item
  consent, \enquote{correct}, \enquote{accuracy}, \enquote{performance}
\item
  draft discussion
\end{itemize}

@@

\subsubsection{\texorpdfstring{Intro to the language (\textbf{mc}) -
please feel free to throw away anything that is not
useful!}{Intro to the language (mc) - please feel free to throw away anything that is not useful!}}\label{intro-to-the-language-mc---please-feel-free-to-throw-away-anything-that-is-not-useful}

\begin{itemize}
\tightlist
\item
  complexity in the vowel system
\item
  complexity in the consonant system
\item
  word shapes
\item
  typical word length
\item
  although not the focus of this paper, high use of suppletion in verbal
  paradigms, other features of language, see Levinson XXX for details
\end{itemize}

@@

A first list of candidate items was generated in 2018 by selecting
simple consonants (\enquote{p}, \enquote{t}, \enquote{d}, \enquote{k},
\enquote{m}, \enquote{n}, \enquote{w}, \enquote{y}) and vowels
(\enquote{i}, \enquote{o},\enquote{u}, \enquote{a},\enquote{e}) that
were combined into consonant-vowel syllables, further sampling the space
of 1- to 4-syllable sequences. These candidates were automatically
checked against Levinson's 2015 dictionary and removed from
consideration if they appeared in the dictionary. The second author
presented them orally to three local research assistants, who were asked
to repeat them and further say whether they were real words. Any item
for which two or more of the assistants reported them having a meaning
or some form of association was excluded.

A second list of candidate items was generated in 2019 by selecting
complex consonants and systematically crossing them with all the vowels
in the Yélî inventory to produce consonant-vowel monosyllables. As
before, items were automatically excluded if they appeared in the
dictionary. Additionally, since hearing vowel length in monosyllables in
isolation is challenging, any item that had a short/long real word
neighbor was filtered out. Since the phonology and phonetics of Yélî is
still in the process of being described {[}CITE \textbf{mc} please fill
in{]}, there could have been undocumented constraints that rendered
items illegal. Therefore, we made sure that the precise consonant-vowel
sequence occurred in some real word in the dictionary (i.e., that there
was a longer word included the monosyllable as a subsequence). These
candidates were presented to one informant, for a final check that they
did not mean anything. Together with the 2018 selection, they were
recorded using a headset \textbf{XXX }mc** please fill in** and an
Olympus \textbf{XXX }mc** please fill in** from the written form
presented together with the same item orally (by the second author). The
complete recorded list was finally presented to two more informants, who
could repeat all the items and who confirmed there were no real words.
Even so, there was one monosyllable that was often identified as a real
word (intended \enquote{yî} /yXX/; identified as \enquote{yi} /yi/,
\emph{tree}). This item is removed from analyses below.

@@

We visited four different hamlets once, and attempted to test all
eligible children present at the time, to prevent the items
\enquote{spreading} through hearsay. In the first hamlet, we tested
children in five different places, with some children being tested
inside a house and others tested on the veranda. The complete list of
places and the ways in which they met the desiderata mentioned above can
be found in the raw data, available from online supplementary materials.

>>>>

\usepackage{fontspec}

%% Pandoc can only set 10, 11, 12 pt
%% uncomment below to set fontsize
%\usepackage[fontsize=13pt]{scrextend}

%\setmainfont{Calibri} % Set main font for latin characters

\setlength{\parskip}{0.15cm} % Set space between paragraphs
%\linespread{1.2}\selectfont % Set line height
%\usepackage[doublespacing]{setspace} % Use double space without changing footnotes line height

%% Special font for IPA
%% Make sure "Doulos SIL" is installed on your computer
%% For other typefaces supporting IPA symbols, see
%% https://en.wikipedia.org/wiki/International_Phonetic_Alphabet#Typefaces
\newfontfamily\ipa{Doulos SIL} % Font for IPA symbols
\DeclareTextFontCommand{\ipatext}{\ipa}

%%%         Section for CJK Characters                   %%%
%%%   You may want to uncomment the code below if        %%%
%%%   you're writing this document with CJK characters   %%%

%\usepackage{xeCJK}  % Uncomment for using CJK characters
%% Set main font for CJK characters
%% Make sure your system has the font set
%\setCJKmainfont[
%	BoldFont={HanWangHeiHeavy}  % Set font for CJK boldface
%    ]{標楷體}    % Set font for normal CJK
%% Some Traditional Chinese fonts: AR PL KaitiM Big5, PingFang TC, Noto Sans CJK TC
%\XeTeXlinebreaklocale "zh"
%\XeTeXlinebreakskip = 0pt plus 1pt

@@

\begin{itemize}
\tightlist
\item
  look through this manuscript for bits of text to be written by
  \textbf{mc}
\item
  help with table 1 (formatting \& making it appear, because it's broken
  by the phonological representations)
\item
  do you think Figure 2 should be done with phonetic symbols? If yes,
  could you try to implement that?
\item
  double check demo -- probably start with the xlsx version that has the
  children's first names, date of birth, and date of test. I looked
  again at my local files (still haven't checked the labs' ones) and I'm
  surprised I have maternal education for more children than I have
  birth order for\ldots{} If you correct this, to generate the csv I
  also round the age, assign age.rounded to reported age (in years) when
  DOB is missing, remove first names, DOB, DOTest and age.precise before
  pushing to github
\item
  double check my frequency entries in segments.xlsx using the search
  function in \url{https://phoible.org/parameters}; take a look at
  \url{http://phoible.github.io/conventions/} in case you see something
  about double articulation (I didn't find tp or lBj)
\item
  double check phonological implementation of stimuli in stimuli.xlsx
\item
  for both of the above, if you make any changes, they need to be
  re-saved as UTF16 txt file, and then converted (see remember above)
\item
  double check also stim\_seg\_freq -- why are some items still with NA
  freq?
\item
  consent, ``correct'', ``accuracy'', ``performance''
\item
  draft discussion
\end{itemize}

@@

\subsubsection{The Yélî language}\label{the-yuxe9luxee-language}

Yélî Dnye is an isolate language (presumed Papuan) spoken by
approximately 7000 people residing on Rossel Island, which is found at
the far end of the Louisiade Archipelago in Milne Bay Province, Papua
New Guinea. Most speakers of the language grow up speaking it
monolingually within their swidden horticulturalist communities until
they begin attending school around the age of 7; school instruction is
in English. While monolingual Yélî Dnye upbringing is common,
multilingual families are not unusual, particularly in the region around
the Catholic Mission (1953--present) where there appears to be a higher
incidence of married-in mothers from other islands ({\textbf{???}}), the
same region in which the current data were collected. Children in these
multilingual families grow up speaking Yélî Dnye and English, Tok Pisin,
and/or some other language(s) from the region.

The Yélî sound system, much like its baroque grammatical system in
general ({\textbf{???}}), is unlike any other in the region, with 90
contrastive segments, including at least two contrasts not yet been
documented elsewhere (labial-coronal double-articulations with dental
vs.~post-alveolar coronal placement in both oral and nasal stops)
({\textbf{???}}). With only four primary places of articulation
(bilabial, alveolar, post-alveolar, and velar) and no voicing contrasts,
the phonological inventory is remarkably packed with acoustically
similar segments. The core oral stop set includes both singleton (/p/,
/t/, /ṭ/, and /k/) and doubly-articulated (/tp/, /ṭp/, /kp/) segments,
with full nasal equivalents (/m/, /n/, /ṇ/, /ŋ/, /nm/, /ṇm/, /ŋm/), and
with a substantial portion of these able to be pre-nasalized or nasally
released (/mp/, /nt/, /ṇṭ/, /ŋk/, /nmtp/, /ṇmṭp/, /ŋmkp/, /ṭṇ/, /kŋ/,
/ṭpṇm/, /kpŋm/). Finally, a number of these can also be labialized or
palatalized on the release, or both (/pʲ/, /tʃ/, /tpʲ/, /ṭʲ/, /ṭpʲ/,
/kʲ/, /kpʲ/, /pʲʷ/, /mbʲ/, /ndʒ/, /nmdbʲ/, /ṇḍʲ/, /ṇmḍbʲ/, /mbʷ/, /ŋɡʷ/,
/mbʲʷ/, /ṭṇʲ/, /ṭpṇmʲ/, /kŋʷ/, /mʲ/, /nʲ/, /nmʲ/, /ṇʲ/, /ṇmʲ/, /mʷ/,
/ŋʷ/, mʲʷ/). Among the remaining consonants are small group of oral
continuants (/w/, /j/, /ɣ/, /l/, /βʲ/, /lʲ/, /lβʲ/). Vowels in Yélî Dnye
may be oral or nasal, short or long. The 10 oral vowel qualities (/i/,
/ɯ/, /u/, /e/, /o/, /ə/, /ɛ/, /\ipatext{ɔ}/, /æ/, /ɑ/) can be produced
as short or long vowels, with 7 of these able to appear as short and
long nasal vowels as well (/i/, /u/, /ə/, /ɛ/, /\ipatext{ɔ}/, /æ/, /ɑ/).
In total, Levinson ({\textbf{???}}) then counts 90 distinctive segments
in Yélî Dnye (93 when including consonants that are extremely rarely
used).

To do: - word shapes - typical word length

@@

A first list of candidate items was generated in 2018 by selecting
simple consonants (``p'', ``t'', ``d'', ``k'', ``m'', ``n'', ``w'',
``y'') and vowels (``i'', ``o'',``u'', ``a'',``e'') that were combined
into consonant-vowel syllables, further sampling the space of 1- to
4-syllable sequences. These candidates were automatically checked
against Levinson's 2015 dictionary and removed from consideration if
they appeared in the dictionary. The second author presented them orally
to three local research assistants, who were asked to repeat them and
further say whether they were real words. Any item for which two or more
of the assistants reported them having a meaning or some form of
association was excluded.

A second list of candidate items was generated in 2019 by selecting
complex consonants and systematically crossing them with all the vowels
in the Yélî inventory to produce consonant-vowel monosyllables. As
before, items were automatically excluded if they appeared in the
dictionary. Additionally, since hearing vowel length in monosyllables in
isolation is challenging, any item that had a short/long real word
neighbor was filtered out. Since the phonology and phonetics of Yélî is
still in the process of being described {[}CITE \textbf{mc} please fill
in{]}, there could have been undocumented constraints that rendered
items illegal. Therefore, we made sure that the precise consonant-vowel
sequence occurred in some real word in the dictionary (i.e., that there
was a longer word included the monosyllable as a subsequence). These
candidates were presented to one informant, for a final check that they
did not mean anything. Together with the 2018 selection, they were
recorded using a headset \textbf{XXX }mc** please fill in** and an
Olympus \textbf{XXX }mc** please fill in** from the written form
presented together with the same item orally (by the second author). The
complete recorded list was finally presented to two more informants, who
could repeat all the items and who confirmed there were no real words.
Even so, there was one monosyllable that was often identified as a real
word (intended ``yî'' /yXX/; identified as ``yi'' /yi/, \emph{tree}).
This item is removed from analyses below.

@@

We visited four different hamlets once, and attempted to test all
eligible children present at the time, to prevent the items `spreading'
through hearsay. In the first hamlet, we tested children in five
different places, with some children being tested inside a house and
others tested on the veranda. The complete list of places and the ways
in which they met the desiderata mentioned above can be found in the raw
data, available from online supplementary materials.
